Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election 

### 2024-04-18 paper/paper/paper(1).qmd
Add files via upload

The Canadian electoral process, epitomizing democratic principles, empowers citizens to elect parliamentarians who shape Parliament's composition and, consequently, the government formation.
In the recent 2021 federal election, Justin Trudeau's Liberal Party secured a majority in the 44th Parliament.
Voters' primary concerns include the cost of living, healthcare, climate change, post-pandemic recovery, and immigration, with dissatisfaction towards Trudeau's government prevalent.
Trudeau's popularity lags behind UK's Poilievre.
The government's ambitious immigration plan aims to admit 500,000 migrants annually by 2025, albeit facing a backlog.
The Conservative Party, with a focus on economic prosperity, emphasizes attracting highly skilled immigrants while addressing concerns like living costs and housing issues[@Canada].

## Hypothesis

Based on our analysis of historical trends, governmental policies, and electoral procedures, our hypothesis aligns with the prevailing expectation that the Conservatives will emerge as the likely winners in the 2025 election.
**We anticipate that the Conservatives will likely secure the highest number of votes, with the Liberals expected to closely follow in second place.** Our research indicates that the vote-counting process in Canada significantly influences the outcome, as voters do not directly elect a specific prime minister candidate.
Initially, all votes are tallied based on first preferences.
If a candidate obtains 50% of the vote, they win.
If not, the candidate with the fewest votes is eliminated[@Canada].
Then, votes from supporters of the eliminated candidate are redistributed to their second-choice candidate.
This iterative redistribution continues until one candidate achieves 50% of the vote.
Hence, our objective is to forecast the election probabilities for the Conservative and Liberal parties and compare which is higher[@Elections].

@@

### Data Summary Measures

```{r}
#| label: tbl-data-summary
#| tbl-cap: Summary of Numerical Variables Across Census and Survey Data
#| echo: false
numeric_summary_df <- data.frame(
  Variable = c("Age in Census Data", "Age in Survey Data", 
               "Liberal Votes in Survey", "Conservative Votes in Survey"),
  Minimum = c(15.0, 18.0, 0.000, 0.000),
  Q1 = c(37.0, 36.0, 0.000, 0.000),
  Median = c(54.0, 53.0, 0.000, 0.000),
  Average = c(52.18, 51.3, 0.267, 0.249),
  Q3 = c(67.0, 66.0, 1.000, 0.000),
  Maximum = c(80.0, 97.0, 1.000, 1.000)
)

knitr::kable(numeric_summary_df)
```

@tbl-data-summary summaries the census and survey data reveals age distributions with the census data capturing a narrower age range (15 to 80 years, with a median of 54) compared to the broader age span in the survey data (18 to 97 years, with a median of 53).
The average ages for the census and survey populations are 52.18 and 51.3 years, respectively.
Voting data from the survey indicate that a quarter of the respondents are inclined towards the Liberal Party, as shown by an average of 0.267, while a slightly lower average of 0.249 suggests Conservative support; both parties have 50% of the population not voting for them, with a median of 0.
The upper bounds for Liberal and Conservative votes are 1, meaning some respondents indicated a vote for these parties, while the third quartile for Conservative votes indicates less overall support compared to the Liberal Party.

@@

The last stage incorporated post-stratification to gauge the distribution of voter preferences across the identified political factions.
This involved segmenting the populace according to our model's predictors and calculating the voting proportions for each party within these segments.
\newpage

@@

-   $y_i$: Binary outcome of voting for the Liberal/Conservative party (1) or not (0)
-   $\pi_i$: Probability of voting for the Liberal/Conservative party
```{=tex}
\begin{align}
y_i | \pi_i &\sim \text{Bernoulli}(\pi_i) \\
\text{logit}(\pi_i) &= \beta_0 + \alpha^{age}_{a[i]} + \alpha^{gender}_{g[i]} + \alpha^{province}_{p[i]} + \alpha^{religion}_{r[i]} + \alpha^{income}_{m[i]} + \alpha^{education}_{e[i]} \\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\alpha_{a}^{age} &\sim \text{Normal}(0, \sigma^{2}_{age}) \quad \text{for } a \text{ in } 1, 2, \ldots, A \\
\alpha_{g}^{gender} &\sim \text{Normal}(0, \sigma^{2}_{gender}) \quad \text{for } g \text{ in } Male,Female \\
\alpha_{p}^{province} &\sim \text{Normal}(0, \sigma^{2}_{province}) \quad \text{for } p \text{ in } BC, AB, MB, NB, NL, NS, ON, PE, QC, SK \\
\alpha_{r}^{religion} &\sim \text{Normal}(0, \sigma^{2}_{religion}) \quad \text{for } r \text{ in } YES,NO \\
\alpha_{m}^{income} &\sim \text{Normal}(0, \sigma^{2}_{income}) \quad \text{for } m \text{ in } Lower Class, Middle Class, Upper Class\\
\alpha_{e}^{education} &\sim \text{Normal}(0, \sigma^{2}_{education}) \quad \text{for } e \text{ in } Low Education, Some Education, High Education \\
\sigma^{2}_{age} &\sim \text{Exponential}(1) \\
\sigma^{2}_{gender} &\sim \text{Exponential}(1) \\
\sigma^{2}_{province} &\sim \text{Exponential}(1) \\
\sigma^{2}_{religion} &\sim \text{Exponential}(1) \\
\sigma^{2}_{income} &\sim \text{Exponential}(1) \\
\sigma^{2}_{education} &\sim \text{Exponential}(1) 
\end{align}
```

## Model justification

## Model Coefficients Interpretation

@@

$$log(\frac{p}{1-p}) = \beta_0+\beta_1x_{age}+\beta_2x_{gender}+\beta_3x_{porvince}+\beta_{4}x_{religion}+\beta_{5}x_{income}+\beta_{6}x_{education}$$

$p$ is the probability of an individual casting their vote for a certain political party (Liberal or Conservative).

@@

```{r echo=FALSE}
liberal_model <- readRDS(here("models", "liberal_model.rds"))
conservative_model <- readRDS(here("models", "conservative_model.rds"))
```

Table2: Summary the residuals for Conservative Party:
```{r echo=FALSE, warning=FALSE}

liberal_residuals <- posterior_predict(liberal_model) - cleaned_survey_data$vote_liberal
conservative_residuals <- posterior_predict(conservative_model) - cleaned_survey_data$vote_conservative

table_summary <- data.frame(
  Party = c("Liberal", "Conservative"),
  Mean_Residual = c(mean(liberal_residuals, na.rm = TRUE), mean(conservative_residuals, na.rm = TRUE)),
  Median_Residual = c(median(liberal_residuals, na.rm = TRUE), median(conservative_residuals, na.rm = TRUE)),
  SD_Residual = c(sd(liberal_residuals, na.rm = TRUE), sd(conservative_residuals, na.rm = TRUE)),
  IQR_Residual = c(IQR(liberal_residuals, na.rm = TRUE), IQR(conservative_residuals, na.rm = TRUE))
)

# Create a table of the data
kable(table_summary, caption = "Summary of Residuals by Party")

@@

Table3: Summary the coefficient for Liberal Party:

```{r echo=FALSE, warning=FALSE}

liberal_coefs <- coef(liberal_model)
conservative_coefs <- coef(conservative_model)

liberal_coefs_df <- data.frame(Estimate = liberal_coefs, row.names = NULL)
conservative_coefs_df <- data.frame(Estimate = conservative_coefs, row.names = NULL)

liberal_coefs_df$Term <- names(liberal_coefs)
conservative_coefs_df$Term <- names(conservative_coefs)

liberal_coefs_df <- liberal_coefs_df[, c("Term", "Estimate")]
conservative_coefs_df <- conservative_coefs_df[, c("Term", "Estimate")]

knitr::kable(liberal_coefs_df, caption = "Coefficients of the Liberal Model", row.names = FALSE)
knitr::kable(conservative_coefs_df, caption = "Coefficients of the Conservative Model", row.names = FALSE)

@@

# Discussion
Recent research highlights that the primary concerns for Canadian voters leading up to the 2025 election include the rising cost of living, housing, healthcare, the economy, and the environment. These issues are pivotal to the electorate, and the political parties’ stances on them are under intense scrutiny. Our analysis, drawn from five different polls, suggests a marginal preference for the Conservative Party over the Liberal Party. This trend is echoed in Barik's 2023 article, which indicates a public perception of the Conservative Party's superior handling of these key issues. Thus, our working hypothesis posits that the Conservative Party holds a slender advantage in the race for the upcoming election[@Barik].

@@

In conclusion, our binary logistic regression approach, supplemented with post-stratification, has yielded a data-driven forecast for the 2025 Canadian Election. Aligning with Barik’s analysis and multiple polls, our findings suggest the Conservative Party is poised for a narrow victory over the Liberal Party, informed by an intricate blend of sociopolitical factors and current Canadian sentiment[@Barik].

@@

@fig-conservative-model performs a posterior predictive check for a conservative model by overlaying observed vote counts against simulated data derived from the model's posterior distribution.
The observed data is denoted by a solid line, while the simulations from the posterior are indicated by lighter lines.
The congruence between these two sets of lines—where the simulated data traces the shape and trend of the observed data closely—suggests that the model has a high level of accuracy in reflecting the observed outcomes.
This indicates a successful fit of the model to the data, affirming the appropriateness of the selected priors and the model's structure.
The spread of the lighter lines around the darker one provides insight into the inherent uncertainty of the model's predictions, which is an expected characteristic of predictive modelling in complex, real-world situations.

```{r fig-liberal-model, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6, fig.cap="Comparison of Posterior and Prior Distributions in the Liberal Model"}

survey_data <- read_csv(here("data", "analysis_data", "cleaned_survey_data.csv"))

@@

print(pp_plot)

@@

@fig-liberal-model portrays the posterior and prior distributions for parameters in a Bayesian liberal model, with a stark divergence in the intercept suggesting significant data-driven adjustments from the model's baseline estimation.
Parameters representing demographic and geographical factors are plotted as colored dots along an axis, indicating median values and the extent of uncertainty.
While the priors and posteriors largely overlap, indicating consistent estimates for many parameters, the intercept's noticeable shift underscores the data's impactful role in refining the model, particularly highlighting the updated understanding of the model's fundamental starting point.

```{r fig-conservative-posterior, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6, fig.cap="Comparison of Posterior and Prior Distributions in the Conservative Model"}
library(bayesplot)

pp_plot <- posterior_vs_prior(conservative_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10),
    axis.title.y = element_text(size = 7, face = "bold"),
    axis.title.x = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 7)
  ) +
  coord_flip() +
  xlab("Density") +
  ylab("Parameters")

print(pp_plot)

>>>>

The Canadian electoral process, epitomizing democratic principles, empowers citizens to elect parliamentarians who shape Parliament's composition and, consequently, the government formation.
In the recent 2021 federal election, Justin Trudeau's Liberal Party secured a majority in the 44th Parliament[@Nakhaie_2006].
Voters' primary concerns include the cost of living, healthcare, climate change, post-pandemic recovery, and immigration, with dissatisfaction towards Trudeau's government prevalent.
Trudeau's popularity lags behind UK's Poilievre[@Nakhaie_2006].
The government's ambitious immigration plan aims to admit 500,000 migrants annually by 2025, albeit facing a backlog.
The Conservative Party, with a focus on economic prosperity, emphasizes attracting highly skilled immigrants while addressing concerns like living costs and housing issues[@Elections_2023].

## Hypothesis
To enhance control and prevention of illegal immigration, the Liberal Party has suggested modifications to the Immigration and Refugee Protection Act. They propose establishing an annual refugee quota of 250, specifically aimed at supporting human rights defenders, journalists, and humanitarian workers seeking safety and resettlement in Canada. Looking ahead, the Liberal Party intends to incrementally increase this refugee quota, aiming for refugees to comprise 15% of all immigrants. Conversely, the share of economic immigrants is planned to decrease to 58%, while the family reunification immigrant ratio will remain constant. These proportions are expected to be maintained over the next two years. Moreover, the Canadian national government recently unveiled an ambitious plan to admit 500,000 migrants annually by 2025, looking for nearly 1.5 million newcomers in Canada within the next three years (Canada's Immigration Levels Plan 2024-2026). Historically, the Liberal government's immigration strategy tends to be more permissive, offering broader entry routes compared to that of the Conservative Party[@Nakhaie_2006].
Based on our analysis of historical trends, governmental policies, and electoral procedures, our hypothesis aligns with the prevailing expectation that the Conservatives will emerge as the likely winners in the 2025 election.
**We anticipate that the Liberals will likely secure the highest number of votes, with the Conservatives expected to closely follow in second place.** Our research indicates that the vote-counting process in Canada significantly influences the outcome, as voters do not directly elect a specific prime minister candidate.
Initially, all votes are tallied based on first preferences.
If a candidate obtains 50% of the vote, they win.
If not, the candidate with the fewest votes is eliminated[@Elections_2023].
Then, votes from supporters of the eliminated candidate are redistributed to their second-choice candidate.
This iterative redistribution continues until one candidate achieves 50% of the vote.
Hence, our objective is to forecast the election probabilities for the Conservative and Liberal parties and compare which is higher[@Elections_2023].

@@

### Data Summary Measures
```{r}
#| label: tbl-data-summary-numeric
#| tbl-cap: Summary of Numerical Variables Across Census and Survey Data
#| echo: false
numeric_summary_df <- data.frame(
  Variable = c("Age in Census Data", "Age in Survey Data", 
               "Liberal Votes in Survey", "Conservative Votes in Survey"),
  Minimum = c(15.0, 18.0, 0.000, 0.000),
  Q1 = c(37.0, 36.0, 0.000, 0.000),
  Median = c(54.0, 53.0, 0.000, 0.000),
  Average = c(52.18, 51.3, 0.267, 0.249),
  Q3 = c(67.0, 66.0, 1.000, 0.000),
  Maximum = c(80.0, 97.0, 1.000, 1.000)
)

knitr::kable(numeric_summary_df)

@tbl-data-summary-numeric summaries the census and survey data reveals age distributions with the census data capturing a narrower age range (15 to 80 years, with a median of 54) compared to the broader age span in the survey data (18 to 97 years, with a median of 53).
The average ages for the census and survey populations are 52.18 and 51.3 years, respectively.
Voting data from the survey indicate that a quarter of the respondents are inclined towards the Liberal Party, as shown by an average of 0.267, while a slightly lower average of 0.249 suggests Conservative support; both parties have 50% of the population not voting for them, with a median of 0.
The upper bounds for Liberal and Conservative votes are 1, meaning some respondents indicated a vote for these parties, while the third quartile for Conservative votes indicates less overall support compared to the Liberal Party.

@@

The last stage incorporated post-stratification to gauge the distribution of voter preferences across the identified political factions.
This involved segmenting the populace according to our model's predictors and calculating the voting proportions for each party within these segments.

@@

-   $y_i$: Binary outcome of voting for the Liberal/Conservative party (1) or not (0)
-   $\pi_i$: Probability of voting for the Liberal/Conservative party

```{=tex}
\begin{align}
y_i | \pi_i &\sim \text{Bernoulli}(\pi_i) \\
\text{logit}(\pi_i) &= \beta_0 + \beta_1 x_{i,\text{age}} + \beta_2 x_{i,\text{gender}} + \beta_3 x_{i,\text{province}} + \beta_4 x_{i,\text{religion}} + \beta_5 x_{i,\text{income}} + \beta_6 x_{i,\text{education}} \\
\beta_0 &\sim \text{Normal}(0, 1) \\
\beta_1, \beta_2, \beta_3, \beta_4, \beta_5, \beta_6 &\sim \text{Normal}(0, 1)
\end{align}

The model, as formulated in the Bayesian logistic regression framework, is designed to examine the impact of various sociodemographic factors on the probability of an individual voting for a specific political party, such as the Liberals or Conservatives in Canada. The probability $p_i$of a particular voting outcome follows a Bernoulli distribution, where each individual's probability of voting for the party is modeled using logistic regression.

## Components of the Model

## Bernoulli Distribution
Each binary outcome $y_i$ (whether an individual votes for the party or not) follows a Bernoulli distribution with success probability $pi_i$, which is the probability that $y_i = 1$.

### Logit Link Function
The logistic function or logit link transforms the linear combination of predictors into a probability (0 to 1). It is defined as $\log(\frac{\pi_i}{1-\pi_i})$, mapping real numbers from the linear predictors to the interval $[0,1]$.

### Predictors and Coefficients
   - $\beta_0$ (Intercept): Represents the baseline log-odds of voting for the party when all predictor variables are at zero or their reference categories.
   - $\beta_1$ (Age): Reflects the influence of age on the probability of voting for the party. A positive $\beta_1$ indicates increasing age is associated with a higher likelihood of voting for the party.
   - $\beta_2$(Gender): Adjusts the probability based on gender, typically comparing males to females if females are the reference category.
   - $\beta_3$(Province): Accounts for regional variations in voting patterns, with coefficients for each province relative to a reference province.
   - $\beta_4$ (Religion): Represents the difference in voting probability between individuals identifying with a religion versus those who do not.
   - $\beta_5$ (Income): Captures the effect of income on voting behavior.
   - $\beta_6$ (Education): Reflects how varying levels of education influence voting decisions.

### Prior Distributions
All coefficients, including the intercept, are assumed to follow a Normal(0, 1) distribution as priors. These priors are weakly informative, centering the expected value of each effect around zero but allowing the data to inform the final posterior distributions. This approach helps in regularizing the estimates, particularly in complex models or when data are limited.

The model effectively quantifies how demographic and socioeconomic factors influence political behavior using Bayesian inference. This method provides a probabilistic framework that not only gives estimates of the effects but also accounts for uncertainty in a coherent way, offering more flexibility and insight than traditional frequentist approaches. By assuming normal priors, you are indicating a prior belief that none of the predictors has an unusually large effect, but you remain open to being informed by the data collected.

@@

$$log(\frac{p}{1-p}) = \beta_0+\beta_1x_{age}+\beta_2x_{gender}+\beta_3x_{porvince}+\beta_{4}x_{religion}+\beta_{5}x_{income}+\beta_{6}x_{education}$$
## Model Coefficients Interpretation
$p$ is the probability of an individual casting their vote for a certain political party (Liberal or Conservative).

@@

```{r , echo=FALSE, message=FALSE, warning=FALSE}
liberal_model <- readRDS(here("models", "liberal_model.rds"))
conservative_model <- readRDS(here("models", "conservative_model.rds"))
```

```{r, echo=FALSE, message=FALSE, warning = FALSE}
#| label: tbl-coef-summary-conservative
#| tbl-cap: Coefficients of the conservative Model
#| echo: false

install.packages("MASS")

library(MASS)
library(broom)
library(knitr)

initial_model <- lm(vote_conservative ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)

step_cons <- stepAIC(initial_model, direction="both", trace = FALSE)  
tidy_cons <- broom::tidy(step_cons, conf.int = TRUE)

tidy_cons_selected <- tidy_cons[, c("term", "estimate", "std.error", "conf.low", "conf.high")]

kable_output <- knitr::kable(tidy_cons_selected, format = "html", caption = "Model Summary for Conservative Vote Prediction with Confidence Intervals")

kable_output

@@

```{r, echo=FALSE, message=FALSE, warning = FALSE}
#| label: tbl-coef-summary-liberal
#| tbl-cap: Coefficients of the Liberal Model
#| echo: false

# Assuming 'vote_liberal' is the column in your 'cleaned_survey_data' that represents voting for liberal
initial_model_lib <- lm(vote_liberal ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)
step_lib <- stepAIC(initial_model_lib, direction="both", trace = FALSE)

tidy_lib <- broom::tidy(step_lib, conf.int = TRUE)

tidy_lib_selected <- tidy_lib[, c("term", "estimate", "std.error", "conf.low", "conf.high")]

kable_output_lib <- knitr::kable(tidy_lib_selected, format = "html", caption = "Model Summary for Liberal Vote Prediction with Confidence Intervals")
kable_output_lib

@tbl-coef-summary-liberal

`

@tbl-coef-summary-conservative

@tbl-ypshat-summary
```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
if (!require(MASS)) {
    install.packages("MASS")
    library(MASS)
} else {
    library(MASS)
}

m_lib <- glm(vote_liberal ~ age + sex + province + education + religion + income_before_tax,
             data = cleaned_survey_data, family = "binomial")

step_lib <- stepAIC(m_lib, direction = "both", trace = FALSE)

```{r, echo=FALSE, message=FALSE, warning=FALSE, include = FALSE}

m_lib <- glm(vote_liberal ~ age+sex+province+education+religion+income_before_tax,
          data=cleaned_survey_data, family = "binomial")

step_lib = stepAIC(m_lib, direction = "both", k = 2)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, include = FALSE}
m_cons <- glm(vote_conservative ~ age+sex+province+education+religion+income_before_tax,
          data=cleaned_survey_data, family = "binomial")

step_cons <- stepAIC(m_cons, direction = "both", k = 2)

```{r, echo=FALSE, message=FALSE, warning=FALSE, include = FALSE}

step_cons <- glm(vote_conservative ~ age+province+education+religion+income_before_tax,
          data=cleaned_survey_data, family = "binomial")
step_cons_counts <- cleaned_survey_data %>%
  group_by(age, province, education, religion, income_before_tax)%>%
  summarise(N_j=n(), .groups = 'keep')

step_cons_counts$estimate <- 
  step_cons %>%
  predict(newdata = step_cons_counts, type = "response")

cons_aic <- step_cons_counts %>%
  mutate(predict_prob_cons = estimate*N_j) %>% 
  ungroup() %>%
  summarise(yPS_cons_aic = sum(predict_prob_cons)/sum(N_j))

cons_counts <- cleaned_survey_data%>%
  group_by(age, sex, province, education, religion, income_before_tax)%>%
  summarise(N_j=n(), .groups = 'keep')

cons_counts$estimate <- 
  m_cons %>%
  predict(newdata = cons_counts, type = "response")

cons_estimate <- cons_counts %>%
  mutate(predict_prob_cons = estimate*N_j) %>% 
  ungroup() %>%
  summarise(yPS_cons = sum(predict_prob_cons)/sum(N_j))

step_lib <- glm(vote_liberal ~ age+province+education+religion+income_before_tax,
          data=cleaned_survey_data, family = "binomial")
step_lib_counts <- cleaned_survey_data %>%
  group_by(age, province, education, religion, income_before_tax)%>%
  summarise(N_j=n(), .groups = 'keep')

step_lib_counts$estimate <- 
  step_lib %>%
  predict(newdata = step_lib_counts, type = "response")

lib_aic <- step_lib_counts %>%
  mutate(predict_prob_lib = estimate*N_j) %>% 
  ungroup() %>%
  summarise(yPS_lib_aic = sum(predict_prob_lib)/sum(N_j))

lib_counts <- cleaned_survey_data %>%
  group_by(age, sex, province, education, religion, income_before_tax)%>%
  summarise(N_j=n(), .groups = 'keep')

lib_counts$estimate <- 
  m_lib %>%
  predict(newdata = lib_counts, type = "response")

lib_estimate <- lib_counts %>%
  mutate(predict_prob_lib = estimate*N_j) %>% 
  ungroup() %>%
  summarise(yPS_lib = sum(predict_prob_lib)/sum(N_j))

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
Party_Name <- c("Conservative", "Liberal")
yPS_primary <- c(cons_estimate,lib_estimate)
yPS_AIC <- c(cons_aic ,lib_aic)

knitr::kable(tibble(Party_Name, yPS_primary, yPS_AIC), caption = "The Hat of yPS For Each Party, Between Primary Model and Final Model")

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)
library(broom)

required_vars <- c("vote_liberal", "age", "sex", "province", "education", "religion", "income_before_tax")
missing_vars <- setdiff(required_vars, names(cleaned_survey_data))

if(length(missing_vars) > 0) {
  stop(paste("The following required variables are missing in the data frame:", paste(missing_vars, collapse=", ")))
}
model_lib <- lm(vote_liberal ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)
model_summary_lib <- broom::tidy(model_lib, conf.int = TRUE)

ggplot(model_summary_lib, aes(x = estimate, y = reorder(term, estimate))) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Estimate", y = "Coefficient", title = "Coefficient Estimates for Liberal Vote Prediction") +
  theme_minimal()

@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)

# Assuming 'cleaned_survey_data' contains 'province' and 'vote_liberal',
# where 'vote_liberal' is a binary (0/1) indicating whether a vote for liberal was cast.

# Calculate proportions and confidence intervals by province for the liberal party
liberal_votes_by_province <- cleaned_survey_data %>%
  group_by(province) %>%
  summarise(
    proportion = mean(vote_liberal, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    .groups = 'drop'  # Drop grouping for further manipulation if needed
  ) %>%
  ungroup()  # Ensure we remove the grouping for the ggplot

# Create the plot for the liberal party
ggplot(liberal_votes_by_province, aes(x = reorder(province, -proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  # Flip the axes to have provinces on the y-axis
  labs(x = "Proportion", y = "Province", title = "Estimated Proportion of Voting for Liberal by Province") +
  theme_minimal()

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)

# Assuming 'cleaned_survey_data' is your dataframe and it contains the variables 'education' and 'vote_liberal'
# where 'vote_liberal' is a binary (0/1) indicating whether a vote for liberal was cast.

liberal_votes_by_education <- cleaned_survey_data %>%
  group_by(education) %>%
  summarise(
    proportion = mean(vote_liberal, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    .groups = 'drop'  # This will automatically ungroup the data (requires dplyr version 1.0.0 or higher)
  ) %>%
  arrange(desc(proportion))

ggplot(liberal_votes_by_education, aes(x = reorder(education, proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  # Flip the axes to have education levels on the y-axis
  labs(x = "Proportion", y = "Education", title = "Proportion of Votes for Liberal by Education Level") +
  theme_minimal()

@@

# Discussion
Recent research highlights that the primary concerns for Canadian voters leading up to the 2025 election include the rising cost of living, housing, healthcare, the economy, and the environment. These issues are pivotal to the electorate, and the political parties’ stances on them are under intense scrutiny. Our analysis, drawn from five different polls, suggests a marginal preference for the Conservative Party over the Liberal Party. This trend is echoed in Barik's 2023 article, which indicates a public perception of the Conservative Party's superior handling of these key issues. Thus, our working hypothesis posits that the Conservative Party holds a slender advantage in the race for the upcoming election[@Barik_2023].

@@

In conclusion, our binary logistic regression approach, supplemented with post-stratification, has yielded a data-driven forecast for the 2025 Canadian Election. Aligning with Barik’s analysis and multiple polls, our findings suggest the Conservative Party is poised for a narrow victory over the Liberal Party, informed by an intricate blend of sociopolitical factors and current Canadian sentiment[@Barik_2023].

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.width=8, fig.height=6}

@@

conservative_model <- stan_glm(
  vote_conservative ~ age + sex + province + education + religion + income_before_tax,
  data = cleaned_survey_data,
  family = "binomial",
  prior = prior_covariates,
  prior_intercept = prior_intercept,
  chains = 4,
  iter = 2000
)

pp_plot_conservative <- posterior_vs_prior(conservative_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10),
    axis.title.y = element_text(size = 7, face = "bold"),
    axis.title.x = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 7)
  ) +
  coord_flip() +
  xlab("Density") +
  ylab("Parameters")

pp_plot_conservative

@@

@fig-liberal-model
\newpage

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

cons_predictions <- predict(step_cons, newdata = step_cons_counts, type = "response", interval = "confidence")
cons_CI <- data.frame(cons_predictions)

cons_CI$lwr <- cons_CI$cons_predictions - 1.96 * sqrt(cons_CI$cons_predictions * (1 - cons_CI$cons_predictions) / nrow(cons_CI))
cons_CI$upr <- cons_CI$cons_predictions + 1.96 * sqrt(cons_CI$cons_predictions * (1 - cons_CI$cons_predictions) / nrow(cons_CI))

cons_CI$observation <- 1:nrow(cons_CI)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

step_lib_counts <- cleaned_survey_data %>%
  group_by(age, sex, province, education, religion, income_before_tax) %>%
  summarise(N_j=n(), .groups = 'drop')  # Use .groups = 'drop' to remove the grouping

lib_predictions <- predict(step_lib, newdata = step_lib_counts, type = "response")
lib_CI <- data.frame(lib_predictions)
lib_CI$lwr <- lib_CI$lib_predictions - 1.96 * sqrt(lib_CI$lib_predictions * (1 - lib_CI$lib_predictions) / step_lib_counts$N_j)
lib_CI$upr <- lib_CI$lib_predictions + 1.96 * sqrt(lib_CI$lib_predictions * (1 - lib_CI$lib_predictions) / step_lib_counts$N_j)

lib_CI$observation <- seq_len(nrow(lib_CI))
